load the distribution
```{r}
#this is something non-gaussian that Jeff made up.  Meant to mimic the
#Verizon data from the textbook chapter 18.  Then take a look at it using
#hist.

load("dat.rda")
hist(dat,breaks=20)
```

## NB: R has some built in functions for doing resampling, bootstraping, etc
And there are more online... I will mention a few of them as we go. We will also write some ourselves but I am not going to make any attempt to make this efficient or elegant. For learning purposes we are going to do things more or less by hand. With a computer.

## make a resample.


```{r}
# what does that mean?

# Draw samples with replacement (ie, they don't have to be unique) from the
# data.

# So we need a bunch of integers between 1 and 1550?  Simplest way to get
# them is the matlab randi command

# example for 10 values from 1 - 10:

randi(10,1,10)

# So the indices (store in a variable called ind) for our resample can be 
# found like this:
ind=randi(1550,1,1550);

# or get matlab to get the size of the sample using numel
ind=randi(numel(dat),1,numel(dat));

# So to get our resample all we need to do is index dat with ind and assign
# the result to a new variable.  Call it dat_RS.
dat_RS=dat(ind);
```

take a look at our resample:
Use hist!

```{r}
plot(pressure)
```
