# Welcome to Rstudio! 
Rstudio is a great place for organizing your work in R and integrating other programs into your workflow. Basically all of the work you complete in R is going to be in Rstudio. Let's get familiar with the layout of Rstudio.

# There are 4 windows in R studio: 
- The upper left window (the one you are reading right now) Is the script section. This section allows you to edit and save R-scripts that you can then run right in R-studio. It also has a bar at the top of it that lets you switch between different scripts like you switch between different tabs on a search engine. 

- The lower left window is the console section. This is the part of R-studio that is actually  running R. You can actually type commands directly into the window if you just want to run something quickly, but we will mostly be typing stuff into scripts and running from there.  

- The upper right window has several tabs, but for today we will just focus on the environment tab. This is where you can view all of the stuff you save into your R session such as variables and dataframes (more on those later).

- The lower right window contains several useful tabs. You can look through and open files with file, look at any plots you make, see what R packages you have loaded, and view help files. 

# What is an R-markdown file?
The script you currently have loaded is a special kind of R script called a markdown file. R-markdown is a powerful tool for integrating R-code into documents and websites, but for now we are just using it as a convenient scripting format. 
Everything you have been reading up to this point has been part of the markdown section of the Rmarkdown file. Now let's actually create a coding section. We are going to create a block of code using three grave accents (this thing ` in the upper left corner) and {r}
```{r}
# everything in this grey box can be run by the R coding language. 
# If we want to put any notes into these coding blocks, we need to turn them into comments 
# by starting the line with a hashtag#. This will tell R to ignore whatever else is in the line. 

# Let's actually make something we can run now and run it by hitting the green arrow at the top of the 
# block
2+2

# You should be able to see the output below you and in the console window. 
```
```{r}
# Hitting the green arrow at the top of the block will run everything in the block, but let's figure out 
# how to run stuff line be line. 
# By hitting command + enter, we can run just one line of code. 
# Run just the 3+6 command below be selecting the line and hitting command+enter.  
20+40
3+6
17+31

# This is how we are usually going to run the code in our blocks. 
```
Ok, now let's learn how to actually do some stuff with R. 

# Defining Variables:
```{r}
#Whenever we calculate or create something in R, we can save what we made as a variable to be used later.
# we do this by specifying a variable name and then saying it is equal to something we make. 
# For example
x=2+2

#We have saved the output of 2+2 in a variable called a, now we can look at the variable using print()
print(x)
```

# Base Functions 1: Common Functions
```{r}
# Ok, so we have done some basic addition, but let's make R do something more interesting. let's look at 
# some of the special functions that are built into R that can do special tasks. 

# we can use c() to create a vector of different objects ( a vector is like a list)
x=c(3,6,2,7)

# We can get R to calculate the mean of our vector
mean(x)

# or we can get it to calculate the standard deviation of the values in our vector.
sd(x)
```

# Base Functions 2: The Help Function and Function Argument Inputs
```{r}
# The single most important function in R is the help() function. This can be used to get 
# information on any of the other functions in R. 

# Try tunning help() on the mean function we used earlier. A window should pop up on the lower right
help(mean)

# even for a simple function like mean, there is a lot of information here. The most important sections 
# are Usage and Arguments. Usage will show you how the function should look when it is typed into the 
# console or script. Usage also shows you what the default values are for each argument. 
# The arguments section explains all of the difference arguments of the function. It explains to you what
# the different arguments do and what those arguments can accept as inputs. This is probably a good time 
# to explain the different ways you can specify argument inputs. 

# Let's create another vector real quick. 
x=c(1,2,4,7,NA,4,3,7,9,5)

# If we want to calculate the mean of our vector while trimming off 10% of the observations and removing 
# the NA value, we can either:

# A) put all of the inputs directly into the function IN ORDER
mean(x,0.1,TRUE)

# B) specify the values we want for each argument in any order we like.
mean(x=x, na.rm=TRUE, trim=0.1)
```

# Packages:
```{r}
# R has some good built in functions, but the true beauty of R is in its open-source packages. There is 
# an R package for basically every problem you can imagine, you just need to figure out what the package 
# is called and how it works.

# There are a couple of steps to using a package. we will go through these steps for the boot 
# package that some of you will use later, but we will not actually do anything with boot yet.

# First you install the package with 
install.packages("boot")

#Then you load the package with
library(boot)
# and you are now ready to use it!

# It is important to note that the help() function also works with functions from packages!
help(boot)

# You might notice that the help file you just brought up is pretty... complicated. 
# Help files often contain extremely technical explanations, and it takes practice to learn how 
# to read them correctly. We will help guide you to the right parts of the help files in the workshop
```

#Dataframes
```{r}
# Dataframes are the main way in which R stores data. We will not use them too much today, but it is 
# important that you learn at least a little about them. 

# Here is a super basic dataframe.
data = data.frame(name=c("Sam","Mike"), height_cm=c(197,155))

#let's take a look at it
View(data)

# We can select one of our columns using dataframe$column_name
data$Mike

# and we can select a row using dataframe[row_number,]
data[1,]
```

#Plotting 
```{r}
# let's go over some of the basics of making plots in R. R has pretty good plotting software built into 
# it, so we are going to be using that today.

# Let's make a couple of vectors
x=c(1,5,8,3,7,3,5,3,6,4,7)
y=c(1,4,7,3,8,9,5,4,3,6,3)

# We can make a histogram using one vector as inputs using hist()
hist(x)

# We can make a scatter plot or a line graph using two vectors with plot()
plot(x,y)

# We can also extract our inputs from a dataframe. Let's use our dataframe to make a barplot.
barplot(data$height_cm, names.arg=data$name)
```
#A note on doing stuff in R:
There are often many different ways to do things in R. Sometimes one way is more or efficient or better than other ways, but it is often the case where multiple different solutions are equally good. If you do something in this workshop that is a little different from how I do it in the answer sheet, you should not worry about it too much as long as your answer still makes sense :) 



