Sampling distributions. 
The sampling distribution of a statistic represents all of the different different sample statistics we could get when we draw a random sample from a population. 
```{r}
# lets create a super basic normal distribution to represent a sampling distribution.
# the mean of our distribution will be 10 and the standard deviation will be 2.5. 
x = seq(0, 20, by = .1)
y = dnorm(x, mean = 10, sd = 2.5)

# now we plot the distribution using plot(). 
plot(x,y,type='l', main="Example of a sampling distribution", ylab='Probability',
     xlab='Sample Statistic')

# Now lets draw a single sample from an identical sampling distribution using the rnorm function
# If you are not familiar with rnorm, try running help(rnorm) and see what comes up. Stare this sample in # a variable called sample
sample = 

#Now lets draw 5000 samples from an idential distribution and see what they look like using the 
#hist() function. Try help(hist) if stuck.
sample = 
```
This shape probably looks familiar to most of you. This is a normal distribution. It is going to come up a lot for reasons we are about to go over. Let's go back to the slides now. 


The t statistic and the t distribution. 
```{r}
#What does our t distribution look like?
x=seq(-5,5,by=0.001)
tdistr=dt(x,19)
plot(x,tdistr,type="l",main="t distribution",xlab="t",ylab="Porobability Density")

# Look at the plot, and then append the code below to your plot to create a rejection region. 
# + polygon(c(x[x>=1.73], max(x), 1.73), c(tdistr[x>=1.73], 0, 0), col="red") #


```
What we just did is sort of like a manual version of conducting a t-test. R has a whole bunch of built in machinery that makes doing t-tests much simpler. Let's actually do a t-test now. 

Let's say the test we are trying to accomplish is to see if the data is
consistent with a population mean of 10.  This will be our H0: mu=10.

What's our Ha?  
  
Let's say the alternative is that mu > 10, This means we will be looking
for our test statistic in a rejection region which is out to the right
side of the distribution, so we will need to set the greater than option.

One Sample T-test. 

```{r}
# First set the sample size we are going to use when we draw our sample. Let's do a sample size of 20 and
#call the variable n
n = 20

# create a sample to use via the rnorm command. Call this variable "sample".
# Give it a mean of 10 and a standard deviation of 2.5
# plot it in a histogram using hist()
sample = 

# Quickly remake the t distribution, by copying the code we used last time. 
x=seq(-5,5,by=0.001)
tdistr=dt(x,19)
plot(x,tdistr,type="l",main="t distribution",xlab="t",ylab="Porobability Density")

# what's our t value and where is it on the distribution?  What do you predict for the result of
# the test?
obs_t=(mean(sample)-10)/(sd(sample)/sqrt(length(sample)))

# R's t test function is called t.test. run help(t.test) to get a feel for the function 


# do the t.test:
# remember, the mean of our null is 10 and we want to a right one-sdided (greater than) test


# There are a lot of outputs, but for our purposes we want to focus on 
# the p-value. 
```
just for funsies:  How often do we make a type I error?  It's
easy to estimate for type I here.  We know mu.  Above. It's 10.

So what's a type I error again?  H0 rejected when it's true.
In our case H0 is true.
let's generate a sample and do the test lot's of times to see how often
we reject H0.
```{r}
N_iter=20000
h_out=0
alpha=0.05
for (i in 1:N_iter){
  x=rnorm(n, mean=10, sd=2.5)
  p=t.test(x,mu=10,alternative="g")[["p.value"]]
  if (p < alpha){
  h_out=h_out+1
  }
  }
# So what's the rate of type I error?

h_out/N_iter

# It should be close to our significance level!  If H0 is true and we
# reject the Null we are out in the tail of distribution by chance.  And
# alpha is the probability of that!

# Ok that's all for now.  Go back to slides.
```
Two sample T-test.
```{r}
# first make 2 large samples using rnorm again. lets say 200 and 400 samples each. 
# Call the variables x and y.The mean of x will be 10 and the mean of y will be 11. 
# set the standard deviation to 2.5 
x=
y=

# Look at them using hist().

# very cute, but let's reduce the number of data points to make this a
# little more messy, I mean realistic. Lets do sample sizes of x = 20 and y = 40 (same sd).

x=
y=

# Look at them using hist().

```

```{r}
# Here I made the variances equal.  In practice if we are using a two
# sample t test and allowing ourselves to assume equal variances then we
# compute what is called the pooled standard deviation.  

sp=((length(x)-1)*sd(x)^2+(length(y)-1)*sd(y)^2)/(length(x)+length(y)-2)
sp=sqrt(sp)

# and then we need to think about our H0 for a moment.

# let's say our H0 is that the population means are equal.  ie mu1=mu2.  So
# we can calculate our t statistic as follows:
  
obs_t=(mean(x)-mean(y))/sqrt((sp^2)/length(x)+(sp^2)/length(y))

# Now we need to think about our Ha.  Let's say we thought it might be the
# case that mu1 < mu2, That means our t stat would be out there in the left
# tail of the distribution.  Think about an extreme case where we would feel confident
# in saying mu1 < mu2.  Imagine mu1 was 0 and
# mu2 was 1000.  If the samples were large and xbar ~mu1 and ybar ~mu2 then
# tstat would be large and negative! This means to the left.

# ok so what does our t distribution look like?  Where does our observed
# value of t land?  What does that imply for our testing?

tdistr=dt(seq(-5,5,by=.1),length(x)+length(y)-2)
plot(seq(-5,5,by=.1),tdistr,type="l",main="t distribution, 2 samp test example",xlab="",ylab="")

# Ok now use the t.test to confirm our suspicions. look through the help file and figure out how to set 
# the variances as equal


# Note the test can also be done without assuming the variances are equal

# In this case the t stat is a bit easier to calculate, just as we had in
# the presentation.  No pooled sd.  However this test is less accurate than 
# a testmade with the assumption of equal variances.

# Due to this approximation the test stat is only approximately t
# distributed so you only get approximately alpha significance.
```
Paired T-test
```{r}
# ok we need two samples again. Let's call them x and y again. 
# set both sample sizes to 200 this time.set the mean of one of x to 10 and the mean of y to 11. 
# standard deviation for both is 2.5

x=
y=

# ok, so what are our diffs? does the order matter? x-y? y-x? Check at the end to see 
# if your results make sense. 
the_diffs=

# What's our H0.  Let's say it is that the means are equal again.

# what's our t stat?

obs_t=mean(the_diffs)/(sd(the_diffs)/sqrt(length(the_diffs)))

# what does our t distribution look like? Where is our observed t value?
# What's the likely outcome of our test?

tdistr=dt(seq(-5,5,by=.1),length(the_diffs)-1)
plot(seq(-5,5,by=.1),tdistr,type="l",main="t distribution, paired test example",xlab="",ylab="")

# What's our Ha?  Let's say that it is mu1<mu2, so we need to look at a
# rejection region in the left tail of the distribution.

# ok do the test.  it's t.test again but this time we are going to specify that they are paired. 
```
